---
title: Comparisons of house mortgage programs
date: 2023.03.21
output: 
  html_document:
    toc: true
    toc_depth: 4
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = F, results = F, comment = F)
```

```{r}
pacman::p_load(Nobitoru, tidyverse, patchwork)
```

## Summary

Principal: 7M period: 10 or 20 years

```{r fig.height=6}

df.summary <- tibble(
  bank = c("Suwa shinkin", "AU jibun", "SBI (variable focus)", "SBI (variable half-year)"),
  principal = 7000000,
  period = 10,
  interest_rate = c(1.84, 0.39, 0.42, 0.65),
  # initial_cost = c(0, 350000, 272000, 173000)
) %>% 
  mutate(initial_cost = case_when(
    bank == "Suwa shinkin" ~ 0,
    bank == "AU jibun" ~ principal * 0.022 + 125000,
    bank == "SBI (variable focus)" ~ principal * 0.022 + 80000+28000+10000,
    bank == "SBI (variable half-year)" ~ 55000 + 80000+28000+10000
  )) %>% 
  group_by(bank) %>% 
  mutate(
    calc_house_mortgage(principal, interest_rate, period, initial_cost, plotit = F) %>% 
      summarise(sum_interest = sum(interest_payment),
                sum_tax_deduction = sum(tax_deduction))
  ) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + initial_cost)) %>% 
  mutate(Plan = "7M for 10 years") %>% 
  bind_rows(
    tibble(
  bank = c("Suwa shinkin", "AU jibun", "SBI (variable focus)", "SBI (variable half-year)"),
  principal = 7000000,
  period = 20,
  interest_rate = c(1.84, 0.39, 0.42, 0.65),
  # initial_cost = c(0, 350000, 272000, 173000)
) %>% 
  mutate(initial_cost = case_when(
    bank == "Suwa shinkin" ~ 0,
    bank == "AU jibun" ~ principal * 0.022 + 125000,
    bank == "SBI (variable focus)" ~ principal * 0.022 + 80000+28000+10000,
    bank == "SBI (variable half-year)" ~ 55000 + 80000+28000+10000
  )) %>% 
  group_by(bank) %>% 
  mutate(
    calc_house_mortgage(principal, interest_rate, period, initial_cost, plotit = F) %>% 
      summarise(sum_interest = sum(interest_payment),
                sum_tax_deduction = sum(tax_deduction))
  ) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + initial_cost)) %>% 
  mutate(Plan = "7M for 20 years") 
  )

df.summary %>% 
  pivot_longer(c(sum_interest, initial_cost, sum_tax_deduction)) %>% 
  mutate(name = factor(name, levels = c("sum_interest", "initial_cost", "sum_tax_deduction"))) %>% 
  mutate(value = ifelse(name == "sum_tax_deduction", value, -1 * value)) %>% 
  ggplot(aes(x = reorder(bank, -profit), y = value)) +
  geom_col(aes(fill = name), width = .5) +
  geom_hline(yintercept = 0, linewidth = 1) +
  geom_label(aes(label = scales::dollar(profit, prefix = "¥"), y = 500000), hjust = 0) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "¥")) +
  scale_fill_manual(values = c("#FFB90F", "#7A8B8B", c("#4EEE94"))) +
  coord_flip(ylim = c(-1200000, 800000)) +
  # hrbrthemes::theme_ipsum_tw(base_size = 13, axis_title_size = 13) +
  theme_bw(13) +
  theme(legend.position = "top") +
  # guides(legend(fill = guide_legend()))
  facet_wrap(~ Plan, ncol = 1) +
  labs(x = NULL, y = "Profit", fill = "Cost") 


```

Principal: 6M period: 10 or 20 years

```{r fig.height=6}

df.summary <- tibble(
  bank = c("Suwa shinkin", "AU jibun", "SBI (variable focus)", "SBI (variable half-year)"),
  principal = 6000000,
  period = 10,
  interest_rate = c(1.84, 0.39, 0.42, 0.65),
  # initial_cost = c(0, 350000, 272000, 173000)
) %>% 
  mutate(initial_cost = case_when(
    bank == "Suwa shinkin" ~ 0,
    bank == "AU jibun" ~ principal * 0.022 + 125000,
    bank == "SBI (variable focus)" ~ principal * 0.022 + 80000+28000+10000,
    bank == "SBI (variable half-year)" ~ 55000 + 80000+28000+10000
  )) %>% 
  group_by(bank) %>% 
  mutate(
    calc_house_mortgage(principal, interest_rate, period, initial_cost, plotit = F) %>% 
      summarise(sum_interest = sum(interest_payment),
                sum_total_payment = sum(total_payment),
                sum_tax_deduction = sum(tax_deduction))
  ) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + initial_cost)) %>% 
  mutate(Plan = "6M for 10 years") %>% 
  bind_rows(
    tibble(
  bank = c("Suwa shinkin", "AU jibun", "SBI (variable focus)", "SBI (variable half-year)"),
  principal = 6000000,
  period = 20,
  interest_rate = c(1.84, 0.39, 0.42, 0.65),
  # initial_cost = c(0, 350000, 272000, 173000)
) %>% 
  mutate(initial_cost = case_when(
    bank == "Suwa shinkin" ~ 0,
    bank == "AU jibun" ~ principal * 0.022 + 125000,
    bank == "SBI (variable focus)" ~ principal * 0.022 + 80000+28000+10000,
    bank == "SBI (variable half-year)" ~ 55000 + 80000+28000+10000
  )) %>% 
  group_by(bank) %>% 
  mutate(
    calc_house_mortgage(principal, interest_rate, period, initial_cost, plotit = F) %>% 
      summarise(sum_interest = sum(interest_payment),
                sum_total_payment = sum(total_payment),
                sum_tax_deduction = sum(tax_deduction))
  ) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + initial_cost)) %>% 
  mutate(Plan = "6M for 20 years") 
  )

df.summary %>% 
  mutate(total_payment_per_month = sum_total_payment / ifelse(str_detect(Plan, "10 years"), 12 * 10, 12 * 20)) %>% 
  pivot_longer(c(sum_interest, initial_cost, sum_tax_deduction)) %>% 
  mutate(name = factor(name, levels = c("sum_interest", "initial_cost", "sum_tax_deduction"))) %>% 
  mutate(value = ifelse(name == "sum_tax_deduction", value, -1 * value)) %>% 
  ggplot(aes(x = reorder(bank, -profit), y = value)) +
  geom_col(aes(fill = name), width = .5) +
  geom_hline(yintercept = 0, linewidth = 1) +
  geom_label(aes(label = scales::dollar(profit, prefix = "¥"), y = 500000), hjust = 0) +
  geom_label(aes(label = scales::dollar(total_payment_per_month, prefix = "¥"), y = 1000000), hjust = 0) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "¥")) +
  scale_fill_manual(values = c("#FFB90F", "#7A8B8B", c("#4EEE94"))) +
  coord_flip(ylim = c(-1200000, 1400000)) +
  # hrbrthemes::theme_ipsum_tw(base_size = 13, axis_title_size = 13) +
  theme_bw(13) +
  theme(legend.position = "top") +
  # guides(legend(fill = guide_legend()))
  facet_wrap(~ Plan, ncol = 1) +
  labs(x = NULL, y = "Profit", fill = "Cost") 


```


Principal: 5M period: 10 or 20 years

```{r fig.height=6}

df.summary <- tibble(
  bank = c("Suwa shinkin", "AU jibun", "SBI (variable focus)", "SBI (variable half-year)"),
  principal = 5000000,
  period = 10,
  interest_rate = c(1.84, 0.39, 0.42, 0.65),
  # initial_cost = c(0, 350000, 272000, 173000)
) %>% 
  mutate(initial_cost = case_when(
    bank == "Suwa shinkin" ~ 0,
    bank == "AU jibun" ~ principal * 0.022 + 125000,
    bank == "SBI (variable focus)" ~ principal * 0.022 + 80000+28000+10000,
    bank == "SBI (variable half-year)" ~ 55000 + 80000+28000+10000
  )) %>% 
  group_by(bank) %>% 
  mutate(
    calc_house_mortgage(principal, interest_rate, period, initial_cost, plotit = F) %>% 
      summarise(sum_interest = sum(interest_payment),
                sum_total_payment = sum(total_payment),
                sum_tax_deduction = sum(tax_deduction))
  ) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + initial_cost)) %>% 
  mutate(Plan = "5M for 10 years") %>% 
  bind_rows(
    tibble(
  bank = c("Suwa shinkin", "AU jibun", "SBI (variable focus)", "SBI (variable half-year)"),
  principal = 5000000,
  period = 20,
  interest_rate = c(1.84, 0.39, 0.42, 0.65),
  # initial_cost = c(0, 350000, 272000, 173000)
) %>% 
  mutate(initial_cost = case_when(
    bank == "Suwa shinkin" ~ 0,
    bank == "AU jibun" ~ principal * 0.022 + 125000,
    bank == "SBI (variable focus)" ~ principal * 0.022 + 80000+28000+10000,
    bank == "SBI (variable half-year)" ~ 55000 + 80000+28000+10000
  )) %>% 
  group_by(bank) %>% 
  mutate(
    calc_house_mortgage(principal, interest_rate, period, initial_cost, plotit = F) %>% 
      summarise(sum_interest = sum(interest_payment),
                sum_total_payment = sum(total_payment),
                sum_tax_deduction = sum(tax_deduction))
  ) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + initial_cost)) %>% 
  mutate(Plan = "5M for 20 years") 
  )

df.summary %>% 
  mutate(total_payment_per_month = sum_total_payment / ifelse(str_detect(Plan, "10 years"), 12 * 10, 12 * 20)) %>% 
  pivot_longer(c(sum_interest, initial_cost, sum_tax_deduction)) %>% 
  mutate(name = factor(name, levels = c("sum_interest", "initial_cost", "sum_tax_deduction"))) %>% 
  mutate(value = ifelse(name == "sum_tax_deduction", value, -1 * value)) %>% 
  ggplot(aes(x = reorder(bank, -profit), y = value)) +
  geom_col(aes(fill = name), width = .5) +
  geom_hline(yintercept = 0, linewidth = 1) +
  geom_label(aes(label = scales::dollar(profit, prefix = "¥"), y = 500000), hjust = 0) +
  geom_label(aes(label = scales::dollar(total_payment_per_month, prefix = "¥"), y = 1000000), hjust = 0) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "¥")) +
  scale_fill_manual(values = c("#FFB90F", "#7A8B8B", c("#4EEE94"))) +
  coord_flip(ylim = c(-1200000, 1400000)) +
  # hrbrthemes::theme_ipsum_tw(base_size = 13, axis_title_size = 13) +
  theme_bw(13) +
  theme(legend.position = "top") +
  # guides(legend(fill = guide_legend()))
  facet_wrap(~ Plan, ncol = 1) +
  labs(x = NULL, y = "Profit", fill = "Cost") 


```




## Half-way full repayment 

5M for 20 years mortgage with AU jibun
```{r}
df <- calc_house_mortgage(5000000, 0.389, 20, 5000000*0.022+125000)

df %>% 
  filter(id <= 120) %>% 
  select(unpaid_balance) %>% 
  tail()

before_10y <- df %>% 
  filter(id <= 120)

before_10y %>% 
  summarise(sum_interest = sum(interest_payment),
            sum_principal_payment = sum(principal_payment),
                sum_total_payment = sum(total_payment),
                sum_tax_deduction = sum(tax_deduction)) %>% 
  mutate(profit = sum_tax_deduction - (sum_interest + 5000000 * 0.022 + 125000)) %>% 
  pull(profit)

# initial cost for AU jibun
5000000 * 0.022 + 125000

# --> 2500040 already paid!
5000000
2645882	/ 120

after_10y <- df %>% 
  filter(id > 120)

```




```{r}

df <- calc_house_mortgage(6000000, 0.39, 20, 6000000*0.022+125000)

df %>% 
  filter(id <= 120) %>% 
  select(unpaid_balance) %>% 
  tail()
```



## Characteristics of banks/plans

-   Suwa shinkin

    -   interest rate: 1.84%

    -   initial cost: 0

    -   Merits:

        -   Quick procedures
        -   No initial cost
        -   Can be used for renovation

    -   Downsides:

        -   High interest rate

-   AU jibun (variable)

    -   interest rate: 0.39%

    -   initial cost: principal x 2.2% + 125,000 = `r 7000000*0.022 + 125000`

-   SBI shinsei

    -   Variable focus
        -   interest rate: 0.42%

        -   initial cost: principal x 2.2% + 118,000 = `r 7000000*0.022 + 118000`
    -   Variable half-year
        -   interest rate: 0.65%

        -   initial cost: 55,000 + 118,000 = `r 55000 + 118000

## 7M for 10 years

### Suwa shinkin

```{r}
calc_house_mortgage(7000000, 1.84, 10, 0)
```

### AU jibun bank

-   interest rate: 0.4%
-   initial cost: 350000

Merits:

-   Low interest rate

Downsides:

-   Not clear how quickly they proceed

```{r}
calc_house_mortgage(7000000, 0.4, 10, 350000)
```

### SBI shinsei

#### Variable focus

-   interest rate: 0.42%
-   initial cost: 272000

Merits:

-   Low interest rate

Downsides:

-   Not clear how quickly they proceed

```{r}
calc_house_mortgage(7000000, 0.42, 10, 272000)
```

#### Variable half-year

-   interest rate: 0.65%
-   initial cost: 173000

Merits:

-   Low interest rate

Downsides:

-   Not clear how quickly they proceed

```{r}
calc_house_mortgage(7000000, 0.65, 10, 173000)
```

## 7M for 20 years

Principal: 7,000,000 Period: 10 years

### Suwa shinkin

-   interest rate: 1.84%
-   initial cost: 0

Merits:

-   Quick procedures

Downsides:

-   High interest rate

```{r}
calc_house_mortgage(7000000, 1.84, 20, 0)
```

### AU jibun bank

-   interest rate: 0.4%
-   initial cost: 350000

Merits:

-   Low interest rate

Downsides:

-   Not clear how quickly they proceed

```{r}
calc_house_mortgage(7000000, 0.4, 20, 350000)
```

### SBI shinsei

#### Variable focus

-   interest rate: 0.42%
-   initial cost: 272000

Merits:

-   Low interest rate

Downsides:

-   Not clear how quickly they proceed

```{r}
calc_house_mortgage(7000000, 0.42, 20, 272000)
```

#### Variable half-year

-   interest rate: 0.65%
-   initial cost: 173000

Merits:

-   Low interest rate

Downsides:

-   Not clear how quickly they proceed

```{r}
calc_house_mortgage(7000000, 0.65, 20, 173000)
```
